DRAFT: Harmonisation and trait aggregation paper

### 2020-12-22 Paper/Draft_harmonization_paper.tex
22.12: Fixing references

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{float}
\usepackage[table,xcdraw]{xcolor}
\usepackage{color, colortbl}
\usepackage{longtable}
%\usepackage[sort&compress,square,comma,authoryear]{natbib}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{rotating}
\usepackage{caption}
% \graphicspath{{/home/kunz/Dokumente/Projects/Trait_DB/Invertebrate_traits/Paper/Figures/}}
\graphicspath{{Figures/}}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{array}
\usepackage{lscape}
\usepackage{longtable}
\usepackage{hhline}
\usepackage{lmodern}
\usepackage[
backend = biber,
style = apa,
citestyle = authoryear,
maxcitenames = 2, 
maxbibnames = 99 % ,
%sorting = none % sort by name year title
]{biblatex}
\addbibresource{Ref_invertebrate_DB.bib}

@@

In the last decades, freshwater ecologists have compiled comprehensive invertebrate trait databases for various continents (\cite{usseglio-polatera_biomonitoring_2000, schmidt-kloiber_www.freshwaterecology.info_2015, vieira_database_nodate, Philips_and_Smith_NZ_DB_2018, kefford_integrated_2020, tomanova_trophic_2006}). The availability of invertebrate trait data from different continents enables comparisons of trait variation and analyses of their relation to environmental factors across large scales. However, such analyses have been carried out mostly within continents, using information from one or two trait databases. For example, Bonada et al. (\cite{bonada_taxonomic_2007}) compared trait composition for Mediterranean and temperate regions in Europe using traits from Usseglio-Polatera et al. (\cite{usseglio-polatera_biomonitoring_2000}, typically referred to as Tachet database); Poff et al. (\cite{poff_developing_2010}) characterized trait composition across sites in the Western US using traits from Poff et al. (\cite{poff_functional_2006}), and Botwe et al. (\cite{botwe_effects_2018}) investigated the effect of salinity on invertebrate traits in different sites in South Australia using trait data from Poff et al. (\cite{poff_functional_2006}) and Schäfer et al. (\cite{schafer_trait_2011}). Analyses of invertebrate traits that synthesize information on invertebrate grouping features from more than two different continents are rare.

In this study, we follow the terminology proposed by Schmera et al. (\cite{schmera_proposed_2015}), where a grouping feature is defined as a general property (e.g. feeding mode) that comprises a "group of related traits (e.g., predator, shredder, etc.) that vary among species or among individuals within a species". Thus, we use the term grouping feature in place of which many studies use the term "trait", and the term trait instead of "trait state", "modality" or "trait category". Traits can be described using different codings (binary, fuzzy) that represent the prevalence of a characteristic in an organism.

To our knowledge, only Brown et al. (\cite{brown_functional_2018}) harmonised grouping features from more than two geographically distant invertebrate trait databases, for a limited set of grouping features (8) and taxa (112), in a study on the influence of decreasing glacier cover on functional diversity and community assembly of invertebrates. We suspect that the heterogeneity of information in freshwater invertebrate trait databases, besides the diversity of taxa across regions, is likely a major reason for the lack of studies across continents. To harmonize grouping features from different regions, first commonly accepted and unambiguous trait definitions are required (\cite{schneider_towards_2019}). In the best case, grouping features would be classified into the same traits across databases or they could easily be harmonised using standardised terminology. However, a lack of standardised terminology of trait definitions and poor metadata quality in many trait databases are common issues throughout the field of trait-based ecology (\cite{baird_toward_2011, kissling_towards_2018}). Secondly, consistent coding of traits facilitates the compatibility of trait data from different databases.

@@

However, invertebrate trait databases are heterogeneous regarding the coding they use for their traits (\cite{culp_incorporating_2011}). For example, in the study from Brown et al. (\cite{brown_functional_2018}) the authors needed to reclassify traits using expert knowledge given that the individual databases employed different coding approaches (i.e. European and New Zealand databases fuzzy coding, North American database binary coding). Databases with standardised and unambiguous traits and a consistent coding of traits would minimize the data processing effort. Discrepancies in the taxonomic resolutions (e.g. species, genus or family) when linking observational taxonomic data to trait database represents another challenge. When observations are on a more precise taxonomic level than data available in the trait databases (e.g. observations on species-level, trait data on genus-level) trait information of the less precise taxonomic level is often assigned (e.g. \cite{szocs_effects_2014, vos_taxonomic_2017}). Conversely, if trait information is only available on more precise taxonomic levels than the observed taxa, traits are aggregated to a less precise taxonomic level (e.g. \cite{poff_functional_2006, szocs_effects_2014, piliere_a._f._h._importance_2016, aspin_extreme_2019}). Studies have used different methods of trait aggregation, e.g. the mean (\cite{magliozzi_functional_2019}), median (\cite{szocs_effects_2014}) or the mode (\cite{piliere_a._f._h._importance_2016}). Up to now, studies on how and to which extent different trait aggregation methods influence trait-based analysis are missing. However, related knowledge would inform future studies regarding the choice of the aggregation method.

The outlined difficulties when working with or synthesizing trait data motivated us to (1) investigate discrepancies in trait definition between trait databases from Europe, North America, Australia, and New Zealand. As a basis for further evaluating different trait aggregation methods and how harmonising definition discrepancies between grouping features can influence trait-environment relationships, we established 4 novel invertebrate trait datasets from these trait databases by harmonising 7 grouping features. We evaluate different trait aggregation methods by (2) comparing aggregated traits with assigned traits by experts at family-level for two of the established trait datasets. (3) we re-analysed data on the effect of anthropogenic salinisation on biological traits by Szöcs et al. (\cite{szocs_effects_2014}) using trait data from the established harmonised European trait datasets and aggregated traits. By comparison with the original analysis, we investigated how harmonising and aggregating trait data can modify the outcome of trait-environment relationships.

@@

We extracted information from 6 trait databases from Europe, North America, Australia, and New Zealand and harmonised 7 grouping features. Trait information for Europe was obtained from the freshwaterecology.info database (\cite{schmidt-kloiber_www.freshwaterecology.info_2015}) and the Tachet database (\cite{usseglio-polatera_biomonitoring_2000}). The freshwaterecology.info contains taxa on species-level, while taxa recorded in the Tachet database are on species, genus and family-level. Trait information for North America was obtained from Twardochleb et al. (\cite{twardochleb_trait_data_2020}) and complemented by Vieira et al. (\cite{vieira_database_nodate}). Data on body form for European and North American taxa was based on expert knowledge (\cite{polatera_personal_information_2020}). For Australia and New Zealand, we used trait databases from Kefford et al. (\cite{kefford_integrated_2020}) and Philips and Smith respectively (\cite{Philips_and_Smith_NZ_DB_2018}). To increase readability we refer to the databases as well as the datasets we extracted from them by the name of the continent they originate from, except for European databases, which we refer to by their commonly used names (freshwaterecology.info and Tachet database).

@@

Aggregated trait affinities of the five trait aggregation methods (\textit{direct\_agg\textsubscript{median}}, \textit{direct\_agg\textsubscript{mean}}, \textit{stepwise\_agg\textsubscript{median}}, \textit{stepwise\_agg\textsubscript{mean}}, and \textit{weighted\_agg}) were compared to trait affinities assigned at family-level by experts, which were available for Australia and North America for a subset of grouping features and taxa. For the Australian dataset, we compared aggregated trait affinities with assigned trait affinities resolved at family-level for the grouping features feeding mode and size using data from Chessman et al. (\cite{chessman_dissolved-oxygen_2018}). In Chessman et al. (\cite{chessman_dissolved-oxygen_2018}) feeding mode is classified similarly as in the Australian dataset except that the trait parasite is missing. We conducted the comparison for the 220 families available in Chessman et al. (\cite{chessman_dissolved-oxygen_2018}). Considering each factor combination of family and trait (in total 8) this amounts to 1760 cases.

For the North American dataset, we compared aggregated trait affinities with assigned trait affinities on family-level for the grouping features feeding mode, respiration, size, voltinism, and locomotion. The assigned trait affinities at family-level are part of the North American database (Twardochleb) and originate from expert knowledge. Trait information was available for 94 families of which all were present in the North American dataset (total number of cases 1598). The traits were on the categorical scale and were converted to binary traits prior to the comparison with aggregated trait affinities.

@@

We repeated the analysis of Szöcs et al. (\cite{szocs_effects_2014}) who studied the effect of anthropogenic salinization on invertebrates in the River Werra in Germany. For the re-analysis we used the established harmonised grouping feature dataset for Europe and additionally aggregated traits using the aforementioned aggregation methods.

The river Werra has been subject to effluents from the potash industry since the mid of the 20th century and allows to study responses of invertebrates and their trait compositions to salinization (\cite{bathe_biological_2011}). Sites downstream, upstream, and close to the salt discharge (transition) were compared regarding their trait composition. Further details can be found in Szöcs et al. (\cite{szocs_effects_2014}).

We substituted 6 of the grouping features from the original data with harmonised grouping features from the European trait dataset. We compared the results of the redundancy analysis (RDA) from the original study to the case when using harmonised grouping features. Specifically, the trait composition expressed as community weighted mean (CWM) traits was ordinated along an electric conductivity gradient. We compared the species scores obtained from the RDA, i.e. the coordinates of the tips of the vectors representing the CWM traits in the bi- or triplots. Following the original study, we identified traits associated with high or low salinity based on their distance to the ordination axis median using the mahalanobis distance. CWM traits with a mahalanobis distance greater than the $97.5 \%$ -quantile of the Chi-square distribution (5.02) were regarded as traits responding to either low or high salinity. For our analysis, we used the same 21 grouping features that Szöcs et al. (\cite{szocs_effects_2014}) used. The 6 harmonised grouping features used were \textit{Size, Feeding mode, Locomotion, Oviposition, Respiration}, and \textit{Voltinism}. Additionally, for testing the effect of aggregated traits we assigned to each taxon in Szöcs et al. (\cite{szocs_effects_2014}) the aggregated trait value for its corresponding family and repeated the RDA.

@@

The original RDA in Szöcs et al. (\cite{szocs_effects_2014}) indicated that downstream sites (high salinity) were characterised by the traits shredder, ovoviviparity, multivoltinism, long life cycle duration ($> 1$ year), and gill respiration and upstream sites (low salinity) were characterized by univoltinism, oviposition in clutches and short life cycle duration ($< 1$ year).

@@

Synthesizing trait information from multiple trait databases is crucial to unfold the full potential of trait-based approaches, e.g. for studying community trait responses to environmental gradients worldwide. Our attempt of harmonising invertebrate trait information from trait databases of different regions showed that grouping feature harmonisation is labour-intensive because grouping features are differentiated differently into traits between trait databases and various codings are used to describe traits. In addition, the same traits are sometimes defined differently in the investigated databases, requiring expert knowledge to reclassify these traits (e.g. the trait piercer in the Tachet database). Although, others have noted the lack of standardised trait terminology in freshwater ecology (\cite{baird_toward_2011, brink_traits-based_2011}), we are the first to describe invertebrate trait definition discrepancies for some commonly used grouping features. To resolve definition discrepancies and facilitate data synthesis in the future, terminological standards are needed. Harmonised definitions and concepts of traits have been developed in the past for other organism groups, e.g. for plants with the \textit{Thesaurus Of Plant characteristics} (TOP) initiative (\cite{garnier_towards_2017}). The core of this harmonisation initiative is to provide standardised trait definitions and to draw connections to synonyms, related terms, and surrounding concepts by linking to other controlled vocabularies or ontologies in the field. Such initiatives could be a role model for freshwater ecologists to establish unambiguous terminologies for invertebrate traits. The existing freshwater invertebrate trait databases could be linked through standardised terminologies or ontologies, as suggested by Baird et al. (\cite{baird_toward_2011}) almost a decade ago. By following the recently proposed \textit{Ecological Trait-data Standard Vocabulary} (ETS) providers of invertebrate trait data could connect their traits to such ontologies (\cite{schneider_towards_2019}). Once standardised trait definitions are established, these will improve trait data sharing, trait data processing when working with multiple trait databases, and ultimately interpretation of the derived results.

@@

When trait values are only available at species-level and observational data are on less precise taxonomic levels various trait aggregation methods have been used by researchers. We aggregated traits from an Australian and North American trait dataset to family-level using 5 different aggregation methods and compared the results to traits assigned at family-level for these regions. Evaluation of the differences between aggregated and assigned traits is difficult because it remains unclear what the true value of a particular trait for a particular family is. Aggregation of trait information on species or genus-level to a point estimate on family-level suggests a precision that is not necessarily present, especially not for traits with a high variability or if trait information on species-level is missing. Some traits can vary strongly on more precise taxonomic levels than the family-level. For example, Monaghan and Soares (\cite{monaghan_improving_2013}) found a high intra-family trait dispersion in the Tachet database for traits of the grouping features body size, flow preference, and reproduction. Furthermore, studies focusing on the lability of traits, i.e. how much traits are unconstrained by phylogeny, found that traits of ecological preferences (e.g. thermal preference), body size, resistance forms, and to a lesser extent feeding mode are labile, and thus possibly highly variable (\cite{poff_functional_2006, wilkes_traitbased_2020}). However, if trait aggregation is necessary our study can give guidance, under which circumstances the choice of aggregation method might be important. The 5 aggregation methods we tested use either the mean or the median and different weightings. Our results indicated that 1) the median aggregation methods compared to mean aggregation methods were often closer to the assigned traits at family-level and 2) that the different weightings tested exerted a minor impact on the aggregated trait affinities. We obtained for both datasets more differing cases for mean aggregation methods, irrespective of the weighting. In the Australian dataset, the amount of differing cases was slightly higher for the mean aggregation methods (23 \%) than for median aggregation methods (16 to 18 \% differing cases). In the North American dataset, mean aggregation methods yielded in 47 \% of the cases a difference to the assigned traits, much higher in comparison to the median aggregation methods (between 15 and 18 \%). When differences occurred between median aggregation methods and assigned traits, these differences were greater compared to the mean aggregation methods, particularly for the North American dataset. There, the mean absolute differences for the median aggregations were twice as high (0.63 - 0.7) compared to the mean aggregations (0.3 - 0.31). The differences can be explained by the binary coding used in the North American trait dataset and in the assigned traits. As a result, traits have affinities of either 1 or 0. Therefore, traits of a particular grouping feature have in most cases a higher difference in trait affinities to each other than if they were fuzzy coded. Using the median to aggregate binary coded taxa will often result in a value of 0 or 1, while using the mean will often result in values between 0 and 1. Thus, using the median led for the North American dataset either to agreement with the assigned traits or to higher differences compared to the mean aggregation methods.

@@

With the increasing use of computer vision in ecology, both species identification and trait quantification could be automated (\cite{milosevic_application_2020, hoye_deep_2020}). This might result in more observational data and trait data on species-level in the future. However, if traits need to be aggregated we showed that trait aggregation is often in conclusion with expert assignments, specifically aggregation approaches using the median. Which aggregation method the objectively better measure is cannot be answered, but when trait aggregation is necessary 1) median aggregation seems to be favourable, 2) highly variable traits and taxa where a few genera have a high number of species should be treated carefully, 3) coding of traits should be considered, and 4) weighting trait information based on the taxa present in the used trait database does not seem to have much influence. No matter which aggregation approach is chosen, for future studies in which researchers must aggregate traits, a measure of trait dispersion should be reported to indicate the uncertainty of the aggregated estimate.

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{float}
\usepackage[table,xcdraw]{xcolor}
\usepackage{color, colortbl}
\usepackage{longtable}
%\usepackage[sort&compress,square,comma,authoryear]{natbib}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{rotating}
\usepackage{caption}
% \graphicspath{{/home/kunz/Dokumente/Projects/Trait_DB/Invertebrate_traits/Paper/Figures/}}
\graphicspath{{Figures/}}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{array}
\usepackage{lscape}
\usepackage{longtable}
\usepackage{hhline}
\usepackage{lmodern}
\usepackage[
backend = biber,
natbib,
style = apa,
citestyle = authoryear,
maxcitenames = 2, 
maxbibnames = 99, 
uniquename=false,
uniquelist=false
%sorting = none % sort by name year title
]{biblatex}
\addbibresource{Ref_invertebrate_DB.bib}

@@

In the last decades, freshwater ecologists have compiled comprehensive invertebrate trait databases for various continents (\cite{usseglio-polatera_biomonitoring_2000, schmidt-kloiber_www.freshwaterecology.info_2015, vieira_database_nodate, Philips_and_Smith_NZ_DB_2018, kefford_integrated_2020, tomanova_trophic_2006}). The availability of invertebrate trait data from different continents enables comparisons of trait variation and analyses of their relation to environmental factors across large scales. However, such analyses have been carried out mostly within continents, using information from one or two trait databases. For example, \citet{bonada_taxonomic_2007} compared trait composition for Mediterranean and temperate regions in Europe using traits from \citet{usseglio-polatera_biomonitoring_2000} (typically referred to as Tachet database); \citet{poff_developing_2010}) characterized trait composition across sites in the Western US using traits from \citet{poff_functional_2006}, and \citet{botwe_effects_2018} investigated the effect of salinity on invertebrate traits in different sites in South Australia using trait data from \citet{poff_functional_2006}) and \citet{schafer_trait_2011}. Analyses of invertebrate traits that synthesize information on invertebrate grouping features from more than two different continents are rare.

In this study, we follow the terminology proposed by \citet{schmera_proposed_2015}, where a grouping feature is defined as a general property (e.g. feeding mode) that comprises a "group of related traits (e.g., predator, shredder, etc.) that vary among species or among individuals within a species". Thus, we use the term grouping feature in place of which many studies use the term "trait", and the term trait instead of "trait state", "modality" or "trait category". Traits can be described using different codings (binary, fuzzy) that represent the prevalence of a characteristic in an organism.

To our knowledge, only \citet{brown_functional_2018} harmonised grouping features from more than two geographically distant invertebrate trait databases, for a limited set of grouping features (8) and taxa (112), in a study on the influence of decreasing glacier cover on functional diversity and community assembly of invertebrates. We suspect that the heterogeneity of information in freshwater invertebrate trait databases, besides the diversity of taxa across regions, is likely a major reason for the lack of studies across continents. To harmonize grouping features from different regions, first commonly accepted and unambiguous trait definitions are required (\cite{schneider_towards_2019}). In the best case, grouping features would be classified into the same traits across databases or they could easily be harmonised using standardised terminology. However, a lack of standardised terminology of trait definitions and poor metadata quality in many trait databases are common issues throughout the field of trait-based ecology (\cite{baird_toward_2011, kissling_towards_2018}). Secondly, consistent coding of traits facilitates the compatibility of trait data from different databases.

@@

However, invertebrate trait databases are heterogeneous regarding the coding they use for their traits (\cite{culp_incorporating_2011}). For example, in the study from \citet{brown_functional_2018} the authors needed to reclassify traits using expert knowledge given that the individual databases employed different coding approaches (i.e. European and New Zealand databases fuzzy coding, North American database binary coding). Databases with standardised and unambiguous traits and a consistent coding of traits would minimize the data processing effort. Discrepancies in the taxonomic resolutions (e.g. species, genus or family) when linking observational taxonomic data to trait database represents another challenge. When observations are on a more precise taxonomic level than data available in the trait databases (e.g. observations on species-level, trait data on genus-level) trait information of the less precise taxonomic level is often assigned (e.g. \cite{szocs_effects_2014, vos_taxonomic_2017}). Conversely, if trait information is only available on more precise taxonomic levels than the observed taxa, traits are aggregated to a less precise taxonomic level (e.g. \cite{poff_functional_2006, szocs_effects_2014, piliere_a._f._h._importance_2016, aspin_extreme_2019}). Studies have used different methods of trait aggregation, e.g. the mean (\cite{magliozzi_functional_2019}), median (\cite{szocs_effects_2014}) or the mode (\cite{piliere_a._f._h._importance_2016}). Up to now, studies on how and to which extent different trait aggregation methods influence trait-based analysis are missing. However, related knowledge would inform future studies regarding the choice of the aggregation method.

The outlined difficulties when working with or synthesizing trait data motivated us to (1) investigate discrepancies in trait definition between trait databases from Europe, North America, Australia, and New Zealand. As a basis for further evaluating different trait aggregation methods and how harmonising definition discrepancies between grouping features can influence trait-environment relationships, we established 4 novel invertebrate trait datasets from these trait databases by harmonising 7 grouping features. We evaluate different trait aggregation methods by (2) comparing aggregated traits with assigned traits by experts at family-level for two of the established trait datasets. (3) we re-analysed data on the effect of anthropogenic salinisation on biological traits by \citet{szocs_effects_2014} using trait data from the established harmonised European trait datasets and aggregated traits. By comparison with the original analysis, we investigated how harmonising and aggregating trait data can modify the outcome of trait-environment relationships.

@@

We extracted information from 6 trait databases from Europe, North America, Australia, and New Zealand and harmonised 7 grouping features. Trait information for Europe was obtained from the freshwaterecology.info database (\cite{schmidt-kloiber_www.freshwaterecology.info_2015}) and the Tachet database (\cite{usseglio-polatera_biomonitoring_2000}). The freshwaterecology.info contains taxa on species-level, while taxa recorded in the Tachet database are on species, genus and family-level. Trait information for North America was obtained from \citet{twardochleb_trait_data_2020} and complemented by \citet{vieira_database_nodate}. Data on body form for European and North American taxa was based on expert knowledge (\cite{polatera_personal_information_2020}). For Australia and New Zealand, we used trait databases from Kefford et al. (\cite{kefford_integrated_2020}) and Philips and Smith respectively (\cite{Philips_and_Smith_NZ_DB_2018}). To increase readability we refer to the databases as well as the datasets we extracted from them by the name of the continent they originate from, except for European databases, which we refer to by their commonly used names (freshwaterecology.info and Tachet database).

@@

Aggregated trait affinities of the five trait aggregation methods (\textit{direct\_agg\textsubscript{median}}, \textit{direct\_agg\textsubscript{mean}}, \textit{stepwise\_agg\textsubscript{median}}, \textit{stepwise\_agg\textsubscript{mean}}, and \textit{weighted\_agg}) were compared to trait affinities assigned at family-level by experts, which were available for Australia and North America for a subset of grouping features and taxa. For the Australian dataset, we compared aggregated trait affinities with assigned trait affinities resolved at family-level for the grouping features feeding mode and size using data from \citet{chessman_dissolved-oxygen_2018}. In \citet{chessman_dissolved-oxygen_2018} feeding mode is classified similarly as in the Australian dataset except that the trait parasite is missing. We conducted the comparison for the 220 families available in \citet{chessman_dissolved-oxygen_2018}. Considering each factor combination of family and trait (in total 8) this amounts to 1760 cases.

For the North American dataset, we compared aggregated trait affinities with assigned trait affinities on family-level for the grouping features feeding mode, respiration, size, voltinism, and locomotion. The assigned trait affinities at family-level are part of the North American database (Twardochleb et al.) and originate from expert knowledge. Trait information was available for 94 families of which all were present in the North American dataset (total number of cases 1598). The traits were on the categorical scale and were converted to binary traits prior to the comparison with aggregated trait affinities.

@@

We repeated the analysis of \citet{szocs_effects_2014} who studied the effect of anthropogenic salinization on invertebrates in the River Werra in Germany. For the re-analysis we used the established harmonised grouping feature dataset for Europe and additionally aggregated traits using the aforementioned aggregation methods.

The river Werra has been subject to effluents from the potash industry since the mid of the 20th century and allows to study responses of invertebrates and their trait compositions to salinization (\cite{bathe_biological_2011}). Sites downstream, upstream, and close to the salt discharge (transition) were compared regarding their trait composition. Further details can be found in \citet{szocs_effects_2014}.

We substituted 6 of the grouping features from the original data with harmonised grouping features from the European trait dataset. We compared the results of the redundancy analysis (RDA) from the original study to the case when using harmonised grouping features. Specifically, the trait composition expressed as community weighted mean (CWM) traits was ordinated along an electric conductivity gradient. We compared the species scores obtained from the RDA, i.e. the coordinates of the tips of the vectors representing the CWM traits in the bi- or triplots. Following the original study, we identified traits associated with high or low salinity based on their distance to the ordination axis median using the mahalanobis distance. CWM traits with a mahalanobis distance greater than the $97.5 \%$ -quantile of the Chi-square distribution (5.02) were regarded as traits responding to either low or high salinity. For our analysis, we used the same 21 grouping features that \citet{szocs_effects_2014} used. The 6 harmonised grouping features used were \textit{Size, Feeding mode, Locomotion, Oviposition, Respiration}, and \textit{Voltinism}. Additionally, for testing the effect of aggregated traits we assigned to each taxon in \citet{szocs_effects_2014} the aggregated trait value for its corresponding family and repeated the RDA.

@@

The original RDA in \citet{szocs_effects_2014} indicated that downstream sites (high salinity) were characterised by the traits shredder, ovoviviparity, multivoltinism, long life cycle duration ($> 1$ year), and gill respiration and upstream sites (low salinity) were characterized by univoltinism, oviposition in clutches and short life cycle duration ($< 1$ year).

@@

Synthesizing trait information from multiple trait databases is crucial to unfold the full potential of trait-based approaches, e.g. for studying community trait responses to environmental gradients worldwide. Our attempt of harmonising invertebrate trait information from trait databases of different regions showed that grouping feature harmonisation is labour-intensive because grouping features are differentiated differently into traits between trait databases and various codings are used to describe traits. In addition, the same traits are sometimes defined differently in the investigated databases, requiring expert knowledge to reclassify these traits (e.g. the trait piercer in the Tachet database). Although, others have noted the lack of standardised trait terminology in freshwater ecology (\cite{baird_toward_2011, brink_traits-based_2011}), we are the first to describe invertebrate trait definition discrepancies for some commonly used grouping features. To resolve definition discrepancies and facilitate data synthesis in the future, terminological standards are needed. Harmonised definitions and concepts of traits have been developed in the past for other organism groups, e.g. for plants with the \textit{Thesaurus Of Plant characteristics} (TOP) initiative (\cite{garnier_towards_2017}). The core of this harmonisation initiative is to provide standardised trait definitions and to draw connections to synonyms, related terms, and surrounding concepts by linking to other controlled vocabularies or ontologies in the field. Such initiatives could be a role model for freshwater ecologists to establish unambiguous terminologies for invertebrate traits. The existing freshwater invertebrate trait databases could be linked through standardised terminologies or ontologies, as suggested by \citet{baird_toward_2011} almost a decade ago. By following the recently proposed \textit{Ecological Trait-data Standard Vocabulary} (ETS) providers of invertebrate trait data could connect their traits to such ontologies (\cite{schneider_towards_2019}). Once standardised trait definitions are established, these will improve trait data sharing, trait data processing when working with multiple trait databases, and ultimately interpretation of the derived results.

@@

When trait values are only available at species-level and observational data are on less precise taxonomic levels various trait aggregation methods have been used by researchers. We aggregated traits from an Australian and North American trait dataset to family-level using 5 different aggregation methods and compared the results to traits assigned at family-level for these regions. Evaluation of the differences between aggregated and assigned traits is difficult because it remains unclear what the true value of a particular trait for a particular family is. Aggregation of trait information on species or genus-level to a point estimate on family-level suggests a precision that is not necessarily present, especially not for traits with a high variability or if trait information on species-level is missing. Some traits can vary strongly on more precise taxonomic levels than the family-level. For example, \citet{monaghan_improving_2013} found a high intra-family trait dispersion in the Tachet database for traits of the grouping features body size, flow preference, and reproduction. Furthermore, studies focusing on the lability of traits, i.e. how much traits are unconstrained by phylogeny, found that traits of ecological preferences (e.g. thermal preference), body size, resistance forms, and to a lesser extent feeding mode are labile, and thus possibly highly variable (\cite{poff_functional_2006, wilkes_traitbased_2020}). However, if trait aggregation is necessary our study can give guidance, under which circumstances the choice of aggregation method might be important. The 5 aggregation methods we tested use either the mean or the median and different weightings. Our results indicated that 1) the median aggregation methods compared to mean aggregation methods were often closer to the assigned traits at family-level and 2) that the different weightings tested exerted a minor impact on the aggregated trait affinities. We obtained for both datasets more differing cases for mean aggregation methods, irrespective of the weighting. In the Australian dataset, the amount of differing cases was slightly higher for the mean aggregation methods (23 \%) than for median aggregation methods (16 to 18 \% differing cases). In the North American dataset, mean aggregation methods yielded in 47 \% of the cases a difference to the assigned traits, much higher in comparison to the median aggregation methods (between 15 and 18 \%). When differences occurred between median aggregation methods and assigned traits, these differences were greater compared to the mean aggregation methods, particularly for the North American dataset. There, the mean absolute differences for the median aggregations were twice as high (0.63 - 0.7) compared to the mean aggregations (0.3 - 0.31). The differences can be explained by the binary coding used in the North American trait dataset and in the assigned traits. As a result, traits have affinities of either 1 or 0. Therefore, traits of a particular grouping feature have in most cases a higher difference in trait affinities to each other than if they were fuzzy coded. Using the median to aggregate binary coded taxa will often result in a value of 0 or 1, while using the mean will often result in values between 0 and 1. Thus, using the median led for the North American dataset either to agreement with the assigned traits or to higher differences compared to the mean aggregation methods.

@@

With the increasing use of computer vision in ecology, both species identification and trait quantification could be automated (\cite{milosevic_application_2020, hoye_deep_2020}). This development might result in more observational data and trait data on species-level in the future. However, if traits need to be aggregated we showed that trait aggregation is often in conclusion with expert assignments, specifically aggregation approaches using the median. Which aggregation method the objectively better measure is cannot be answered, but when trait aggregation is necessary 1) median aggregation seems to be favourable, 2) highly variable traits and taxa where a few genera have a high number of species should be treated carefully, 3) coding of traits should be considered, and 4) weighting trait information based on the taxa present in the used trait database does not seem to have much influence. No matter which aggregation approach is chosen, for future studies in which researchers must aggregate traits, a measure of trait dispersion should be reported to indicate the uncertainty of the aggregated estimate.
